Impact of COVID-19 Closures on Revenue and Employment More Severe for Minority-Owned and Dine-In Restaurants

### 2021-02-25 outputs/paper/data_simulation.Rmd
refine data_simulation document. change title to something more specific.

```

@@

First, we read in the skeleton for the survey data. This is what we would hypothetically know about a restaurant before they took the survey. We keep only the ID numbers, the group, and the last 3 characters of the postal code, which the survey-taker will have to enter to verify identity. All other variables, like name and full address, are only kept in the only frame for surveying, which contains no private information.

```{r, warning = FALSE}
# read in the data
# NOTE: scripts 01_scrape_health_depts, 
# 02_stratified_cluster_sampling,
# 03_clean_restaurant_list_csvs, 
# 04_restaurant_survey_frame,
# and 05_import_yelp_data must have been run already

table_for_surveys <- read_csv(here::here("outputs/data/table_for_surveys.csv"))

@@

survey_1
```

@@

First, some research:

```{r, warning = FALSE}
# Read in data
yelp_data <- read_csv(here("inputs/data/yelp_restaurants_ontario.csv")) %>% 
  na.omit() # Remove all NAs
total_restaurants <- count(yelp_data)

# Check proportion of RestaurantsTakeOut
takeout <- yelp_data %>% 
  select(RestaurantsTakeOut)
summary(takeout)
```
```{r, warning = FALSE}
takeout_percentage <- takeout %>%
  group_by(RestaurantsTakeOut) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

takeout_percentage
```

```{r, warning = FALSE}
# Check proportion of RestaurantsTableService
tableserv <- yelp_data %>% 
  select(RestaurantsTableService)
summary(tableserv)
```

```{r, warning = FALSE}
tableserv_percentage <- tableserv %>%
  group_by(RestaurantsTableService) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

tableserv_percentage
```

```{r, warning = FALSE}
# Calculate invalid responses (neither dine-in nor takeout)
invalid <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == FALSE) %>% 
  count()

@@

 service_percentage
```

@@

```{r, warning = FALSE}
# randomly assign service_provided using ratios above

@@

survey_1
```

```{r, warning = FALSE}
# randomly assign service_provided using ratios above

@@

survey_2
```

@@

```{r, warning = FALSE}
# randomly assign disability status using ratios above

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(disability = sample(c("yes", "no", "nonanswer"),
                             size = num_total,
                             prob = c(0.005, .985, 0.01),
                             replace = TRUE))

survey_1
```

```{r, warning = FALSE}
# randomly assign disability status using ratios above

@@

survey_2
```

@@

Some research for racial and ethnic minorities and women in Canada and in business:

```{r, warning = FALSE}
### Load census dataset by Health Regions ###
### Download link: 
### https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/
### download-telecharger/comp/GetFile.cfm?Lang=E&FILETYPE=CSV&GEONO=058
census_2016 <- read.csv(here("inputs", "data", "census_2016.csv"))
census_2016 <- janitor::clean_names(census_2016)

@@

populations

write_csv(populations, here("outputs/data/populations"))
```

```{r, warning = FALSE}
# Split the population info into our treatment and control groups.

@@

populations_split
```

```{r, warning = FALSE}
### Get percentage proportions of demographic groups of interest

@@

populations_split_percentage
```

@@

## Survey 1
```{r, warning = FALSE}
# Randomly assign racial and ethnic minorities, and women, using the ratios above:

@@

survey_1
```

```{r, warning = FALSE}
# Randomly assign racial and ethnic minorities, and women, using the ratios above:

@@

survey_2
```

@@

```{r, warning = FALSE}
# Randomly assign shutdown using the ratios above only for the treatment group:

@@

survey_2
```

@@

From an analysis of Kaggle restaurant data [@kagglerestaurantdata], we can see that restaurant revenue looks like an F distribution.
The distribution rf(n, df1, df2, ncp) = rf(num_rest, 10, 5) gives an approximate shape. Using the f-distribution formula, we know the distribution has a mean of 10/9, so we would multiply by 62,963 to get the mean to 69,959.

@@

```{r, warning = FALSE}
# Randomly sample from the scaled F distribution to generate restaurant revenues:

@@

survey_1
```

@@

```{r, warning = FALSE}
# Randomly sample from the scaled F distribution to generate restaurant revenues, 
# and apply a random number around 1 to add variation:

@@

survey_2
```

@@

$82,000 in yearly sales (gross revenue) per full-time equivalent employee [@restfact]

Therefore 6,833 monthly revenue per full-time equivalent employee.

In January 2021, there were 2187300 full time sales and service employees in Canada and 1164000 part time sales and service employees [@hourswages].
Therefore there are 1.9 full time employees per part time employee.
Part time employees in service and sales averaged 16.6 hours per week. We will assume that 2 PTE = 1 FTE, or 1 PTE = 0.5 FTE.
Therefore, for every 1.9 + 1 employees, there are 1.9 + 1*0.5 FTE. For every 2.9 employees, there are 2.4 FTE.

@@

```{r, warning = FALSE}
# Use revenue to generate realistic numbers of employees for restaurants:

@@

survey_1
```

@@

For the second survey, we will assume that treatment restaurants shifted towards part time random amounts-- that is, the number of full time employees per part time employee dropped from 1.9. X is some number less than or equal to 1.9.

```{r, warning = FALSE}
# Use revenue to generate realistic numbers of employees for restaurants:

@@

survey_2

```

@@

```{r, warning = FALSE}
# Randomly select only a subset of all the generated responses to keep, 
# simulating nonresponse. Also randomly select columns to delete in partial responses, 
# and delete the data from those rows/columns:

@@

survey_1
```

```{r, warning = FALSE}
# Randomly select only a subset of all the generated responses to keep, 
# simulating nonresponse. Also randomly select columns to delete in partial responses, 
# and delete the data from those rows/columns:

@@

survey_2
```

```{r, warning = FALSE}
# Save the two survey results tables:

@@

```{r}
# research for typical budget numbers
restaurants = 2006
survey_total = 2
# https://www.canadapost.ca/cpc/en/business/postal-services/mailing/letter-discounts.page?
indiv_postage = 0.85 
# https://www.ontario.ca/document/your-guide-employment-standards-act-0/minimum-wage
minimum_wage = 14.25 
# https://www.alchemer.com/plans-pricing/small-teams/
management_service = 1300 
average_call_mins = 7
response_rate = .3

@@

write_csv(survey_budget, here("outputs/data/survey_budget.csv"))
```

>>>>

# read in the data
# NOTE: scripts 01_scrape_health_depts, 
# 02_stratified_cluster_sampling,
# 03_clean_restaurant_list_csvs, 
# 04_restaurant_survey_frame,
# and 05_import_yelp_data must have been run already
```

@@

First, we read in the skeleton for the survey data in Table \@ref(tab:start). This is what we would hypothetically know about a restaurant before they took the survey. We keep only the ID numbers, the group, and the last 3 characters of the postal code, which the survey-taker will have to enter to verify identity. All other variables, like name and full address, are only kept in the only frame for surveying, which contains no private information.

```{r start, echo = FALSE, message = FALSE, warning = FALSE}
table_for_surveys <- read_csv(here::here("outputs/data/table_for_surveys.csv"))

@@

survey_1 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Frame for surveys") %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

@@

First, we used Yelp restaurant data to see how many Ontario restaurants offered different types of service in Table \@ref(tab:yelptakeout) and Table \@ref(tab:yelptable) [@yelp_dataset].

```{r, include = FALSE}
# Read in data
yelp_data <- read_csv(here("inputs/data/yelp_restaurants_ontario.csv")) %>% 
  na.omit() # Remove all NAs
total_restaurants <- count(yelp_data)

# Check proportion of RestaurantsTakeOut
takeout <- yelp_data %>% 
  select(RestaurantsTakeOut)
summary(takeout)
```

```{r yelptakeout, echo = FALSE, message = FALSE, warning = FALSE}
takeout_percentage <- takeout %>%
  group_by(RestaurantsTakeOut) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

takeout_percentage %>%
  knitr::kable(booktabs = TRUE, caption = "Yelp Ontario take-out data") %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

```{r, include = FALSE}
# Check proportion of RestaurantsTableService
tableserv <- yelp_data %>% 
  select(RestaurantsTableService)
summary(tableserv)
```

```{r yelptable, echo = FALSE, message = FALSE, warning = FALSE}
tableserv_percentage <- tableserv %>%
  group_by(RestaurantsTableService) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count/sum(count))

tableserv_percentage %>%
  knitr::kable(booktabs = TRUE, caption = "Yelp Ontario table service data") %>%
  kableExtra::kable_styling() 
```

We used these percentages to estimate proportions of restaurants with take-out and dine-in service, in Table \@ref(tab:serviceprop).

```{r serviceprop, echo = FALSE, message = FALSE, warning = FALSE}
# Calculate invalid responses (neither dine-in nor takeout)
invalid <- yelp_data %>% 
  filter(RestaurantsTableService == FALSE & RestaurantsTakeOut == FALSE) %>% 
  count()

@@

 service_percentage %>%
  knitr::kable(booktabs = TRUE, caption = "Estimated service type proportions") %>%
  kableExtra::kable_styling()
```

@@

## Survey 1 and 2

We randomly sampled to assign each restaurant to be "dinein", "both", or "takeout", with probabilities of 0.06, 0.615, and 0.325, in Table \@ref(tab:servicetype1) and Table \@ref(tab:servicetype2).

```{r servicetype1, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign service_provided using ratios above

@@

survey_1 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding service type to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r servicetype2, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign service_provided using ratios above

@@

survey_2 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding service type to Survey 2 data") %>%
  kableExtra::kable_styling()
```

@@

## Survey 1 and 2

We randomly sampled to assign each restaurant to either "yes", "no", or "nonanswer", with probabilities of 0.005, 0.985, and 0.01, in Table \@ref(tab:disability1) and Table \@ref(tab:disability2).

```{r disability1, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign disability status using ratios above

set.seed(19893)
survey_1 <- survey_1 %>%
  mutate(disability = sample(c("yes", "no", "nonanswer"),
                             size = num_total,
                             prob = c(0.005, 0.985, 0.01),
                             replace = TRUE))

survey_1 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner disability status to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r disability2, echo = FALSE, message = FALSE, warning = FALSE}
# randomly assign disability status using ratios above

@@

survey_2 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner disability status to Survey 2 data") %>%
  kableExtra::kable_styling()
```

@@

We researched racial and ethnic minorities and women using census data broken down by local health authorities [@census_2016], shown in Table \@ref(tab:LHApop), Table \@ref(tab:groupscount), and Table \@ref(tab:groupsprop).

```{r LHApop, echo = FALSE, message = FALSE, warning = FALSE}
### Load census dataset by Health Regions ###
### Download link: 
### https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/
### download-telecharger/comp/GetFile.cfm?Lang=E&FILETYPE=CSV&GEONO=058
census_2016 <- read.csv(here("inputs", "data", "census_2016.csv"))
census_2016 <- janitor::clean_names(census_2016)

@@

populations %>%
  slice(1,2,3,5) %>%
  select(-Ontario, -total_treat, -total_control) %>%
  mutate(Info = c("Total Population", "Indigenous", "Visible Minority", "Women")) %>%
  knitr::kable(booktabs = TRUE, caption = "Populations in Ontario local health authorities") %>%
  kableExtra::kable_styling(latex_options = "scale_down")

#save this table
write_csv(populations, here("outputs/data/populations"))
```

```{r groupscount, echo = FALSE, message = FALSE, warning = FALSE}
# Split the population info into our treatment and control groups.

@@

populations_split %>%
  slice(1,2,3,5) %>%
  knitr::kable(booktabs = TRUE, caption = "Populations in treatment and control groups") %>%
  kableExtra::kable_styling()
```

```{r groupsprop, echo = FALSE, message = FALSE, warning = FALSE}
### Get percentage proportions of demographic groups of interest

@@

populations_split_percentage %>%
  slice(1,2,3,5) %>%
  knitr::kable(booktabs = TRUE, caption = "Proportions in treatment and control groups") %>%
  kableExtra::kable_styling()
```

@@

## Survey 1 and 2

We randomly sampled to assign each restaurant to either "yes", "no", or "nonanswer", for the owner being Indigenous, a visible minority, or a woman, using the proportions above (with 0.01 set aside for "nonresponse"), in Table \@ref(tab:minority1) and Table \@ref(tab:minority2).

```{r minority1, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly assign racial and ethnic minorities, and women, using the ratios above:

@@

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner minority status and gender to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r minority2, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly assign racial and ethnic minorities, and women, using the ratios above:

@@

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding owner minority status and gender to Survey 2 data") %>%
  kableExtra::kable_styling()
```

@@

We randomly sampled to assign each restaurant to either "none", "temporary", or "permanent" for shutdowns, with probabilities of 0.558, 0.432, and 0.01, in Table \@ref(tab:closures).

```{r closures, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly assign shutdown using the ratios above only for the treatment group:

@@

survey_2 %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding closures to Survey 2 data") %>%
  kableExtra::kable_styling()
```

@@

From an analysis of Kaggle restaurant data [@kagglerestaurantdata], we can see that restaurant revenue looks like an F-distribution.
The distribution rf(n, df1, df2, ncp) = rf(num_rest, 20, 20) gives an approximately correct shape. Using the F-distribution formula, we know the distribution has a mean of 10/9, so we would multiply by 62,963 to get the mean to 69,959.

@@

We randomly sampled from the above F-distribution to generate revenues for the restaurants, in Table 
\@ref(tab:revenue1).

```{r revenue1, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly sample from the scaled F distribution to generate restaurant revenues:

@@

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding revenue to Survey 1 data") %>%
  kableExtra::kable_styling()
```

@@

We started by randomly sampling from the same F-distribution multiplied by 62963, but also multiplied by random sampling from normal distribution centred at 1, with a standard deviation of 0.1, to add some variation from the previous data. For the treatment group data, we then applied a factor that reduced revenue based on level of shutdown (0.5 for temporary and 0.25 for permanent), another factor that reduced revenue based on service type (0.55 for dine-in and 0.83 for both), and another factor that reduced revenue based on minority status (0.88 for minority business owners). This can be seen in Table \@ref(tab:revenue2).

```{r revenue2, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly sample from the scaled F distribution to generate restaurant revenues, 
# and apply a random number around 1 to add variation:

@@

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding revenue to Survey 2 data") %>%
  kableExtra::kable_styling()
```

@@

Restaurants averaged $82,000 in yearly sales (gross revenue) per full-time equivalent employee [@restfact]. Therefore, we can estimate 6,833 monthly revenue per full-time equivalent employee.

In January 2021, there were 2187300 full time sales and service employees in Canada and 1164000 part time sales and service employees [@hourswages]. Therefore, there are 1.9 full time employees per part time employee.
Part time employees in service and sales averaged 16.6 hours per week [@hourswages]. We will assume that 2 PTE = 1 FTE, or 1 PTE = 0.5 FTE.
Therefore, for every 1.9 + 1 employees, there are 1.9 + 1*0.5 FTE. For every 2.9 employees, there are 2.4 FTE.

@@

We divided revenue by 6833, multiplied by 0.79 and rounded to find the number of full time employees, and multiplied by 0.42 and rounded to find the number of part time employees, in Table \@ref(tab:employ1).

```{r employ1, echo = FALSE, message = FALSE, warning = FALSE}
# Use revenue to generate realistic numbers of employees for restaurants:

@@

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding employment numbers to Survey 1 data") %>%
  kableExtra::kable_styling()
```

@@

For the second survey, we will assume that treatment restaurants shifted towards part time random amounts-- that is, the number of full time employees per part time employee dropped from 1.9, a small drop that varied per restaurant.

We divided revenue by 6833. We used the absolute value of random sample of a normal distribution centred at zero, with a standard deviation of 0.1, as a factor to decrease the part-time ratio of 1.9.
We then used the new ratio to calculate the number of full-time and part-time employees. This can be seen in Table \@ref(tab:employ2).

```{r employ2, echo = FALSE, message = FALSE, warning = FALSE}
# Use revenue to generate realistic numbers of employees for restaurants:

@@

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding employment numbers to Survey 2 data") %>%
  kableExtra::kable_styling(latex_options = "scale_down")
```

@@

## Survey 1 and 2

We randomly sampled to assign each row as "yes", "partial", or "no", with probabilities 0.6, 0.01, and 0.39, for nonresponse. We then discarded all the "no" rows. For the partial rows, we randomly selected a column to delete the data for. This can be seen in Table \@ref(tab:response1) and Table \@ref(tab:response2).

```{r response1, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly select only a subset of all the generated responses to keep, 
# simulating nonresponse. Also randomly select columns to delete in partial responses, 
# and delete the data from those rows/columns:

@@

survey_1 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding nonresponse to Survey 1 data") %>%
  kableExtra::kable_styling()
```

```{r response2, echo = FALSE, message = FALSE, warning = FALSE}
# Randomly select only a subset of all the generated responses to keep, 
# simulating nonresponse. Also randomly select columns to delete in partial responses, 
# and delete the data from those rows/columns:

@@

survey_2 %>%
  arrange(ID) %>%
  slice(101:110) %>%
  knitr::kable(booktabs = TRUE, linesep ='', caption = "Adding nonresponse to Survey 2 data") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

```{r, include = FALSE}
# Save the two survey results tables:

@@

We Calculated our budget based on Canada Post's business costs [@canadapost], Ontario's minimum wage [@minimumwage], and pricing for data management systems for surveys [@alchemer]. The budget can be seen in Table \@ref(tab:budget).

```{r budget, echo = FALSE, message = FALSE, warning = FALSE}
# research for typical budget numbers
restaurants = 2006
survey_total = 2
# https://www.canadapost.ca/cpc/en/business/postal-services/mailing/letter-discounts.page?
indiv_postage = 0.85 
# https://www.ontario.ca/document/your-guide-employment-standards-act-0/minimum-wage
minimum_wage = 14.25 
# https://www.alchemer.com/plans-pricing/small-teams/
management_service = 1300 
average_call_mins = 7
response_rate = .3

@@

survey_budget %>%
  knitr::kable(booktabs = TRUE, caption = "Total survey budget") %>%
  kableExtra::kable_styling(latex_options = "hold_position")

#save the survey for presentation in the main report.
write_csv(survey_budget, here("outputs/data/survey_budget.csv"))
```

\clearpage

### 2021-02-25 outputs/paper/paper.Rmd
refine data_simulation document. change title to something more specific.

---
title: "COVID-19 Closures Impact Restaurant Revenue and Employment Unequally"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "September 20th, 2021"
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "A randomized controlled trial to identify the impact of mandatory partial closures on Ontario restaurants"
abstract: "This report covers an experiment conducted by Petit Poll to test the impact of COVID-19 closures on Ontario restaurants. A randomized controlled trial was conducted and surveys were used to collect results. The data show that restaurant revenue and employee counts were negatively impacted by a mandatory two-week partial closure, and dine-in restaurants and restaurants owned by visible minorites were more negatively impacted. These results suggest that if more closures are necessary, Ontario should assist restaurant owners by offering support for take-out services and minority-owned businesses."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---

>>>>

---
title: "Impact of COVID-19 Closures on Revenue and Employment More Severe for Minority-Owned and Dine-In Restaurants"
author: "Lorena Almaraz De La Garza, Amy Farrow, and Kumalasari Sondjaja"
date: "September 20th, 2021"
output:
  bookdown::pdf_document2:
    toc: yes
subtitle: "A randomized controlled trial of mandatory partial closures and Ontario restaurants"
abstract: "This report covers an experiment conducted by Petit Poll to test the impact of COVID-19 closures on Ontario restaurants. A randomized controlled trial was conducted and surveys were used to collect results. The data show that restaurant revenue and employee counts were negatively impacted by a mandatory two-week partial closure, and dine-in restaurants and restaurants owned by visible minorites were more negatively impacted. These results suggest that if more closures are necessary, Ontario should assist restaurant owners by offering support for take-out services and minority-owned businesses."
thanks: 'Code and data are available at: [github.com/amycfarrow/ontariorestaurantclosuresexperiment](https://github.com/amycfarrow/ontariorestaurantclosuresexperiment).'
bibliography: references.bib
---
